---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Exploration of 2016-2017 CITES Poaching Database"
subtitle: "https://github.com/emilyrwood/CITES_Poaching_Wood"
author: "Emily Wood"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r, include=FALSE}
# Set your working directory

getwd()

# Load your packages

library('formatR')
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)
library(rworldmap)


# Set your ggplot theme

mynewtheme <- theme_grey(base_size = 12) +
  theme(axis.text = element_text(color = "Dark Green"), 
        legend.position = "top")

theme_set(mynewtheme)

# Load your datasets

CITESPoaching2016 <- read_csv("/home/guest/EDA_2022/EDA-Fall2022/Project/comptab_2018-01-29 16_00_comma_separated.csv")

colnames(CITESPoaching2016)
```


```{r, include=FALSE}
#Processing 
#Goal:  data cleaning, integration, transformation, and reduction


# Replace codes with labels for 'Purpose" column. 

CITESPoaching2016$Purpose <- ifelse (CITESPoaching2016$Purpose=="B", "Breeding",
                             ifelse ( CITESPoaching2016$Purpose=="E", "Educational",
                             ifelse ( CITESPoaching2016$Purpose=="G", "Garden",
                             ifelse ( CITESPoaching2016$Purpose=="H", "Hunting",
                             ifelse ( CITESPoaching2016$Purpose=="L", "Law",
                             ifelse ( CITESPoaching2016$Purpose=="M", "Medical",
                             ifelse ( CITESPoaching2016$Purpose=="R", "Reintroduction",
                             ifelse ( CITESPoaching2016$Purpose=="P", "Personal",
                             ifelse ( CITESPoaching2016$Purpose=="Q", "Circus",
                             ifelse ( CITESPoaching2016$Purpose=="S", "Scientific",
                             ifelse ( CITESPoaching2016$Purpose=="T", "Commercial",
                             ifelse ( CITESPoaching2016$Purpose=="Z", "Zoo",
                             "Unknown"))))))))))))

#rename some of the columns

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Importer reported quantity")] <- "Importer_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Exporter reported quantity")] <- "Exporter_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "App.")] <- "Appendix"

```

```{r}
#wrangling 
#Goal: cleaning the raw dataset into a format compatible

CITESPoaching2016$Reported_quantity <- CITESPoaching2016$Importer_reported_quantity + CITESPoaching2016$Exporter_reported_quantity 

#Limit dataset to poaching occurrence measured by count not by weight. 
#I'm choosing count because the majority of cases are recorded in count not weight. 
#Select relevant columns 

CITESPoaching2016["Unit"][is.na(CITESPoaching2016["Unit"])] <- "Count"

CITES2016_Processed<-CITESPoaching2016 %>%
                    filter(Unit == "Count") %>%
                    select(Year, Appendix, Class, Genus, 
                           Importer, Exporter, Importer_reported_quantity, 
                           Exporter_reported_quantity, Term, Purpose, Reported_quantity)  %>%
                     group_by(Appendix)

#Save the processed dataframe to processed folder 
getwd()

write.csv(CITES2016_Processed, row.names = FALSE, file = "./Data_Processed/CITES2016_Processed.csv")

```


```{r}

Export.districution <- ggplot(CITESPoaching2016, aes(x=Exporter_reported_quantity ))+ geom_histogram(binwidth=50, colour="blue", fill="light blue", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Export Distribution") 
print(Export.districution)

Import.distribution <- ggplot(CITESPoaching2016, aes(x=Importer_reported_quantity ))+ geom_histogram(binwidth=50, colour="dark green", fill="light green", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Import Distribution") 
print(Import.distribution)

Graphnew <- 
  ggplot(CITESPoaching2016, aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = Appendix, alpha = .3))+
  geom_point() +
  xlim(0, 4000000) +
  ylim(0, 4000000)

print(Graphnew) 

Appendix1 <- 
  ggplot(subset(CITESPoaching2016, Appendix == "I"), aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = Appendix, alpha = .3)) +
  geom_point() 
print(Appendix1)

#Remove Null Values and combine Export and Import amounts into a total reported quantity column. Previous graphs should not contain the zeros.  

CITESPoaching2016["Importer_reported_quantity"][is.na(CITESPoaching2016["Importer_reported_quantity"])] <- 0

CITESPoaching2016["Exporter_reported_quantity"][is.na(CITESPoaching2016["Exporter_reported_quantity"])] <- 0


#Visualizing countries with greater number of exports and imports
#Found journal to visualize countries https://journal.r-project.org/archive/2011-1/RJournal_2011-1_South.pdf
#install.packages('rworldmap')
library(rworldmap)

#Importer Countries
CITES_Importer_withmap <- joinCountryData2Map( CITES2016_Processed
                             ,joinCode = "ISO_A2"
                             ,nameJoinColumn = "Importer")
mapDevice() #create world map shaped window
mapCountryData(CITES_Importer_withmap, nameColumnToPlot='Importer')



Importerplot1 <- ggplot(CITES2016_Processed, aes(x=Importer))+ 
                geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Importerplot1)

#Difficult to interpret. Wrangle new dataset and create bar graph with only top 10 import countries:

table(CITES2016_Processed$Importer)

ImporterData <-CITES2016_Processed %>%
                group_by(Importer)  %>%
                summarise(Importer, count=n()) %>%
                filter(Importer == "US" | Importer == "JP" | Importer == "DE" | Importer == "FR"
                       | Importer == "HK" | Importer == "CH" | Importer == "CN" | Importer == "SG" | Importer == "AE" | Importer == "CA")


Importerplot2 <- ggplot(ImporterData, aes(x=Importer))+ 
  geom_bar(binwidth=50, colour="purple", fill="lavender", alpha=0.3)
print(Importerplot2)
  
#Exporter Countries 
CITES_Exporter_withmap <- joinCountryData2Map( CITES2016_Processed
                                               ,joinCode = "ISO_A2"
                                               ,nameJoinColumn = "Exporter")
mapDevice() #create world map shaped window
mapCountryData(CITES_Exporter_withmap, nameColumnToPlot='Exporter')

Exporterplot1 <- ggplot(CITES2016_Processed, aes(x=Exporter))+ 
  geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Exporterplot1)

#Difficult to interpret. Wrangle new dataset and create bar graph with only top 10 export countries:

table(CITES2016_Processed$Exporter)

ExporterData <-CITES2016_Processed %>%
  group_by(Exporter)  %>%
  summarise(Exporter, count=n()) %>%
  filter(Exporter == "NL" | Exporter == "ID" | Exporter == "IT" | Exporter == "US"
         | Exporter == "FR" | Exporter == "DE" | Exporter == "EC" | Exporter == "AU" | Exporter == "TH"| Exporter== "ZA")


Exporterplot2 <- ggplot(ExporterData, aes(x=Exporter))+ 
  geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Exporterplot2)

#Which Classes get poached most often? 

Classplot1 <- ggplot(CITES2016_Processed, aes(x=Class ))+ 
  geom_bar(aes(fill=Class))+
  labs(title = "Classes of poaching occurances 2016",
       x = "Class",
       y = "Count") +
  theme(legend.position = "none")+
  coord_flip()

print(Classplot1)


#This has a lot of NAs perhaps they didn't record always the Class 
#Faceted by appendix 

Classplot2 <- ggplot(CITES2016_Processed, aes(x=Class ))+ 
  geom_bar(aes(fill=Class))+
  labs(title = "Classes of poaching occurances 2016",
       x = "Class",
       y = "Count") +
  coord_flip()+
  facet_grid(rows = "Appendix")+
  theme(axis.text = element_text(size = 5))
print(Classplot2)

#create a facet dataset for Appendix I

AppendixI_Facet <- CITES2016_Processed %>%
                    filter(Appendix == "I")


#A closer look at Mammalia 

Mammalia_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Mammalia"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
      legend.key.size = unit(.2, 'cm'),
      legend.key.height = unit(.2, 'cm'), 
      legend.key.width = unit(.2, 'cm'),
      legend.title = element_text(size=3), 
      legend.text = element_text(size=3))

print(Mammalia_terms)

#A closer look at Aves 

Aves_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Aves"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(Aves_terms)

#A closer look at Reptilia

Reptilia_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Reptilia"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(Reptilia_terms)

#Why arent plants represented in Class? monocotyledons or dicotyledons not depicted so potentiall in NAs

AppendixI_Facet["Class"][is.na(AppendixI_Facet["Class"])] <- "Not_recorded"
 
NA_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Not_recorded"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(NA_terms)

```


# Rationale and Research Questions

Does level of protection, class, term affect the amount exported/imported?

What might the best set of predictors for Poaching?


```{r}

#Two way ANOVA


#Linear modeling 

#1.How Does Reported Poaching Quantity vary over Class and Term for only Appendix 1.

Poaching.2way <- aov(data = AppendixI_Facet, Reported_quantity ~ Class + Term)
summary(Poaching.2way)

#P-value is small. Reject our null hypothesis that the mean across our groups is different. 
#Not enough information to tell us which groups are differing

Poaching.2way2lm <- lm(data = AppendixI_Facet, Reported_quantity ~ Class + Term)
summary(Poaching.2way2lm)

#means are different. Reject null 

TukeyHSD(Poaching.2way)

#check interaction between the variables 

Poaching.2way3 <- aov(data = CITES2016_Processed, Reported_quantity ~ Class * Appendix)
summary(Poaching.2way3)

#Interaction is not significant between variables




#2. How Does Reported Poaching Quantity vary over Class and Appendix for entire data

Poaching.2way.test <- aov(data = CITES2016_Processed, Reported_quantity ~ Class + Appendix)
summary(Poaching.2way.test)

#P-value is small. Reject our null hypothesis that the mean across our groups is different. 
#Not enough information to tell us which groups are differing

Poaching.2way2lm.test <- lm(data = CITES2016_Processed, Reported_quantity ~ Class + Appendix)
summary(Poaching.2way2lm.test)

#means are different. Reject null 

TukeyHSD(Poaching.2way.test)

#check interaction between the variables 

Poaching.2way3 <- aov(data = CITES2016_Processed, Reported_quantity ~ Class * Appendix)
summary(Poaching.2way3)

#significant because the pvalue is <2e-16 

TukeyHSD(Poaching.2way3) 

Poaching.interaction <- with(CITES2016_Processed, interaction(Class, Appendix))

Poaching.anova.2way5 <- aov(data = CITES2016_Processed, Reported_quantity ~ Poaching.interaction)

Poaching.groups <- HSD.test(Poaching.anova.2way5, "Poaching.interaction", group = TRUE)
Poaching.groups

Poaching.anova.plot <- ggplot(CITES2016_Processed, aes(y = Reported_quantity, x = Class, color = Appendix )) +
  geom_boxplot()
print(Poaching.anova.plot)

```


\newpage

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
