---
title: "Exploration of 2016-2017 CITES Poaching Database"
author: "Emily Wood"
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
subtitle: https://github.com/emilyrwood/CITES_Poaching_Wood
geometry: margin=2.54cm
fontsize: 12pt
mainfont: Times New Roman
---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r, include=FALSE}
# Set your working directory

getwd()

# Load your packages

library('formatR')
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)
library(rworldmap)


# Set your ggplot theme

mynewtheme <- theme_grey(base_size = 12) +
  theme(axis.text = element_text(color = "Dark Green"), 
        legend.position = "top")

theme_set(mynewtheme)

# Load your datasets

CITESPoaching2016 <- read_csv("/home/guest/EDA_2022/EDA-Fall2022/Project/comptab_2018-01-29 16_00_comma_separated.csv")

colnames(CITESPoaching2016)
```


```{r, include=FALSE}
#Processing 
#Goal:  data cleaning, integration, transformation, and reduction


# Replace codes with labels for 'Purpose" column. 

CITESPoaching2016$Purpose <- ifelse (CITESPoaching2016$Purpose=="B", "Breeding",
                             ifelse ( CITESPoaching2016$Purpose=="E", "Educational",
                             ifelse ( CITESPoaching2016$Purpose=="G", "Garden",
                             ifelse ( CITESPoaching2016$Purpose=="H", "Hunting",
                             ifelse ( CITESPoaching2016$Purpose=="L", "Law",
                             ifelse ( CITESPoaching2016$Purpose=="M", "Medical",
                             ifelse ( CITESPoaching2016$Purpose=="R", "Reintroduction",
                             ifelse ( CITESPoaching2016$Purpose=="P", "Personal",
                             ifelse ( CITESPoaching2016$Purpose=="Q", "Circus",
                             ifelse ( CITESPoaching2016$Purpose=="S", "Scientific",
                             ifelse ( CITESPoaching2016$Purpose=="T", "Commercial",
                             ifelse ( CITESPoaching2016$Purpose=="Z", "Zoo",
                             "Unknown"))))))))))))

#rename some of the columns

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Importer reported quantity")] <- "Importer_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Exporter reported quantity")] <- "Exporter_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "App.")] <- "Appendix"

```

```{r}
#wrangling 
#Goal: cleaning the raw dataset into a format compatible

CITESPoaching2016$Reported_quantity <- CITESPoaching2016$Importer_reported_quantity + CITESPoaching2016$Exporter_reported_quantity 

#Limit dataset to poaching occurrence measured by count not by weight. 
#I'm choosing count because the majority of cases are recorded in count not weight. 
#Select relevant columns 

CITESPoaching2016["Unit"][is.na(CITESPoaching2016["Unit"])] <- "Count"

CITES2016_Processed<-CITESPoaching2016 %>%
                    filter(Unit == "Count") %>%
                    select(Year, Appendix, Class, Genus, 
                           Importer, Exporter, Importer_reported_quantity, 
                           Exporter_reported_quantity, Term, Purpose, Reported_quantity)  %>%
                     group_by(Appendix)

#Save the processed dataframe to processed folder 
getwd()

write.csv(CITES2016_Processed, row.names = FALSE, file = "./Data_Processed/CITES2016_Processed.csv")

```



```{r}

Export.districution <- ggplot(CITESPoaching2016, aes(x=Exporter_reported_quantity ))+ geom_histogram(binwidth=50, colour="blue", fill="light blue", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Export Distribution") 
print(Export.districution)

Import.distribution <- ggplot(CITESPoaching2016, aes(x=Importer_reported_quantity ))+ geom_histogram(binwidth=50, colour="dark green", fill="light green", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Import Distribution") 
print(Import.distribution)

Graphnew <- 
  ggplot(CITESPoaching2016, aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = Appendix, alpha = .3))+
  geom_point() +
  xlim(0, 4000000) +
  ylim(0, 4000000)

print(Graphnew) 

Appendix1 <- 
  ggplot(subset(CITESPoaching2016, Appendix == "I"), aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = Appendix, alpha = .3)) +
  geom_point() 
print(Appendix1)

#Remove Null Values and combine Export and Import amounts into a total reported quantity column. Previous graphs should not contain the zeros.  

CITESPoaching2016["Importer_reported_quantity"][is.na(CITESPoaching2016["Importer_reported_quantity"])] <- 0

CITESPoaching2016["Exporter_reported_quantity"][is.na(CITESPoaching2016["Exporter_reported_quantity"])] <- 0


#Visualizing countries with greater number of exports and imports
#Found journal to visualize countries https://journal.r-project.org/archive/2011-1/RJournal_2011-1_South.pdf
#install.packages('rworldmap')
library(rworldmap)

#Importer Countries
CITES_Importer_withmap <- joinCountryData2Map( CITES2016_Processed
                             ,joinCode = "ISO_A2"
                             ,nameJoinColumn = "Importer")
mapDevice() #create world map shaped window
mapCountryData(CITES_Importer_withmap, nameColumnToPlot='Importer')



Importerplot1 <- ggplot(CITES2016_Processed, aes(x=Importer))+ 
                geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Importerplot1)

#Difficult to interpret. Wrangle new dataset and create bar graph with only top 10 import countries:

table(CITES2016_Processed$Importer)

ImporterData <-CITES2016_Processed %>%
                group_by(Importer)  %>%
                summarise(Importer, count=n()) %>%
                filter(Importer == "US" | Importer == "JP" | Importer == "DE" | Importer == "FR"
                       | Importer == "HK" | Importer == "CH" | Importer == "CN" | Importer == "SG" | Importer == "AE" | Importer == "CA")


Importerplot2 <- ggplot(ImporterData, aes(x=Importer))+ 
  geom_bar(binwidth=50, colour="purple", fill="lavender", alpha=0.3)
print(Importerplot2)
  
#Exporter Countries 
CITES_Exporter_withmap <- joinCountryData2Map( CITES2016_Processed
                                               ,joinCode = "ISO_A2"
                                               ,nameJoinColumn = "Exporter")
mapDevice() #create world map shaped window
mapCountryData(CITES_Exporter_withmap, nameColumnToPlot='Exporter')

Exporterplot1 <- ggplot(CITES2016_Processed, aes(x=Exporter))+ 
  geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Exporterplot1)

#Difficult to interpret. Wrangle new dataset and create bar graph with only top 10 export countries:

table(CITES2016_Processed$Exporter)

ExporterData <-CITES2016_Processed %>%
  group_by(Exporter)  %>%
  summarise(Exporter, count=n()) %>%
  filter(Exporter == "NL" | Exporter == "ID" | Exporter == "IT" | Exporter == "US"
         | Exporter == "FR" | Exporter == "DE" | Exporter == "EC" | Exporter == "AU" | Exporter == "TH"| Exporter== "ZA")


Exporterplot2 <- ggplot(ExporterData, aes(x=Exporter))+ 
  geom_bar(binwidth=50, colour="blue", fill="light blue", alpha=0.3)
print(Exporterplot2)

#Which Classes get poached most often? 

Classplot1 <- ggplot(CITES2016_Processed, aes(x=Class ))+ 
  geom_bar(aes(fill=Class))+
  labs(title = "Classes of poaching occurances 2016",
       x = "Class",
       y = "Count") +
  theme(legend.position = "none")+
  coord_flip()

print(Classplot1)


#This has a lot of NAs perhaps they didn't record always the Class 
#Faceted by appendix 

Classplot2 <- ggplot(CITES2016_Processed, aes(x=Class ))+ 
  geom_bar(aes(fill=Class))+
  labs(title = "Classes of poaching occurances 2016",
       x = "Class",
       y = "Count") +
  coord_flip()+
  facet_grid(rows = "Appendix")+
  theme(axis.text = element_text(size = 5))
print(Classplot2)

#create a facet dataset for Appendix I

AppendixI_Facet <- CITES2016_Processed %>%
                    filter(Appendix == "I")


#A closer look at Mammalia 

Mammalia_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Mammalia"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
      legend.key.size = unit(.2, 'cm'),
      legend.key.height = unit(.2, 'cm'), 
      legend.key.width = unit(.2, 'cm'),
      legend.title = element_text(size=3), 
      legend.text = element_text(size=3))

print(Mammalia_terms)

#A closer look at Aves 

Aves_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Aves"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(Aves_terms)

#A closer look at Reptilia

Reptilia_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Reptilia"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(Reptilia_terms)

#Why arent plants represented in Class? monocotyledons or dicotyledons not depicted so potentiall in NAs

AppendixI_Facet["Class"][is.na(AppendixI_Facet["Class"])] <- "Not_recorded"
 
NA_terms <- ggplot(subset(AppendixI_Facet, Class %in% "Not_recorded"), aes(x = Term)) + 
  geom_bar(aes(fill= Term)) +
  coord_flip() +
  theme(axis.text = element_text(size = 5),
        legend.key.size = unit(.2, 'cm'),
        legend.key.height = unit(.2, 'cm'), 
        legend.key.width = unit(.2, 'cm'),
        legend.title = element_text(size=3), 
        legend.text = element_text(size=3))

print(NA_terms)

```


# Rationale and Research Questions

Poaching remains a considerable threat to biodiversity across the planet. However, it often is overshadowed by other environmental issues in the policy realm. This is especially true for species that aren't charismatic mega-fauna. I'm undertaking this research project to better understand the trends in poaching over time and understand the regions where the most risk is occuring. This could help target poaching hotspots and help policy makers make more informed decisions when it comes to mitigating poaching risk for sesitive species. 

Main Questions:

Is there any correlation between class, term, appendix number and poaching amount?

Sub Questions:

What is the distribetion of amounts found exported and imported for all species included in the database?

Are there differences between Appendix Levels for the amount exported vs. imported?

What are the most commonly poached classes between Appendix Levels?

What are the top 3 Classes for species poached in Appendix I?

What are the dominant good types (terms) for the top 3 classes poached in Appendix I?

Which Counties are the biggest Importer and Exporters of CITES species?


```{r}
#Two way ANOVA

#Linear modeling 

#1.How Does Reported Poaching Quantity vary over Class and Term for only Appendix 1.

Poaching.2way <- aov(data = AppendixI_Facet, Reported_quantity ~ Class + Term)
summary(Poaching.2way)

#P-value is small. Reject our null hypothesis that the mean across our groups is different. 
#Not enough information to tell us which groups are differing

Poaching.2way2lm <- lm(data = AppendixI_Facet, Reported_quantity ~ Class + Term)
summary(Poaching.2way2lm)

#means are different. Reject null 

TukeyHSD(Poaching.2way)

#check interaction between the variables 

Poaching.2way3 <- aov(data = CITES2016_Processed, Reported_quantity ~ Class * Appendix)
summary(Poaching.2way3)

#Interaction is not significant between variables




#2. How Does Reported Poaching Quantity vary over Class and Appendix for entire data

Poaching.2way.test <- aov(data = CITES2016_Processed, Reported_quantity ~ Class + Appendix)
summary(Poaching.2way.test)

#P-value is small. Reject our null hypothesis that the mean across our groups is different. 
#Not enough information to tell us which groups are differing

Poaching.2way2lm.test <- lm(data = CITES2016_Processed, Reported_quantity ~ Class + Appendix)
summary(Poaching.2way2lm.test)

#means are different. Reject null 

TukeyHSD(Poaching.2way.test)

#check interaction between the variables 

Poaching.2way3 <- aov(data = CITES2016_Processed, Reported_quantity ~ Class * Appendix)
summary(Poaching.2way3)

#significant because the pvalue is <2e-16 

TukeyHSD(Poaching.2way3) 

Poaching.interaction <- with(CITES2016_Processed, interaction(Class, Appendix))

Poaching.anova.2way5 <- aov(data = CITES2016_Processed, Reported_quantity ~ Poaching.interaction)

Poaching.groups <- HSD.test(Poaching.anova.2way5, "Poaching.interaction", group = TRUE)
Poaching.groups

Poaching.anova.plot <- ggplot(CITES2016_Processed, aes(y = Reported_quantity, x = Class, color = Appendix )) +
  geom_boxplot()
print(Poaching.anova.plot)

```


\newpage

# Dataset Information

The Convention on International Trade in Endangered Species of Wild Fauna and Flora (CITES) is an international agreement between governments. This group studies poaching trends and advises the UN on necessary action to mitigate this risk. This CITES dataset contains records on every international import or export conducted with species from the CITES monitoring list in 2016 and 2017. It contains columns identifying the species, the import and export countries, and the amount and characteristics of the goods being traded (which range from live animals to skins and cadavers). The 'Term' colum described the type of good the poached species takes when encountered. Examples for this column include live, trophies, tusks, ect. 

Notably, this dataset has a column named, 'Appendix' which sorts the occurrences into three groups. Appendix I species are those whose poaching directly threatens them with extinction whose trade threatens them with extinction. There are around 1200 such species in this Appendix. Appendix II species are those who do not directly face extinction from poaching, but would experience detrimental impacts regardless. This group makes up the majority of the data set with around 21000 species.  Appendix III animals are considered controls because their export and import requires permits. There are around 170 in this dataset. I will be strongly focused on Appendix I in my analysis.

Wrangling the Dataset:

To prepare this dataset I first replaced the Codes for the 'Purpose' column with text to match the other columns such as 'Term'. I then renamed the columns to get rid of abbreviations and spaces. Next I set the NAs in both the importer and exporter quantity columns to 0 so that they could be combined into a total reported quantity column. After this column was created I noted that the Unit column was separate and combined individual occurrence counts and weight in grams and kilograms. After viewing the amounts of each type of character, I made the decision to remove occurrences measured in weight because there were significantly less when compared to those measured in counts. To do this I set The NAs in the Unit column to "Count" and filtered for it when finishing my processed dataframe. As part of this process I also selected the variables I needed which included Year, Appendix, Class, Genus, Importer, Exporter, Importer_reported_quantity, Exporter_reported_quantity, Term, Purpose, Reported_quantity, and I grouped the dataframe by Appendix number. 


| Variables                  | Units        | Ranges                                   |
|----------------------------|--------------|------------------------------------------|
| Year                       |              | 2016-2017                                |
| Appendix                   |              | I,II,III                                 |
| Class                      |              | Actinopteri, Anthozoa, Aves,... Reptilia |
| Genus                      |              | Alligator, Aloe, Alveopora, ...Ursus     |
| Importer                   | Country Code | AD, AE, AU,... ZW                        |
| Exporter                   | Country Code | AD, AE, AR,... ZZ                        |
| Importer_reported_quantity | Count        | 1,2,3, ... 19524978                      |
| Exporter_reported_quantity | Count        | 1,2,3, ... 21543618                      |
| Term                       |              | baleen, bark, bodies, ... wood product   |
| Purpose                    |              | Breeding, Circus, Commercial, ... Zoo    |
| Reported Quantity          | Count        | 1, 2, 3, ... 21543639                    |

               

\newpage

# Exploratory Analysis 




\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
